### 2013-12-02 paper/parts/intro.tex
Began drafting an outline of the introduction.

\section{Introduction}

Lorem ipsum.

\cite{schreiber2000measuring}, \cite{ver2012information}, \cite{shalizi2007discovering}, \cite{paninski2003estimation}

Our SFI paper: \cite{darmon2013detecting}