### 2014-01-13 paper/parts/results.tex
Added more details to the paper about weights across the community boundaries.

\section{Results}

\subsection{Community Statistics (? A better title should go here.)}

Discuss the basic properties of the various types of communities: distribution of sizes, number of singletons, etc.

\begin{figure}[h!]
  \centering
\includegraphics[width=0.50\textwidth]{Figures/overlap_by_type.pdf}
\caption{Lorem ipsum.}
\label{Fig-}
\end{figure}

\begin{figure}[h!]
  \centering
\includegraphics[width=0.50\textwidth]{Figures/comm_sizes_ecdf_loglog.pdf}
\caption{The proportion of communities greater than $c$ in size, across the different community types. Note the logarithmic scale on the horizontal and vertical axes.}
\label{Fig-community_size_distribution}
\end{figure}

\subsection{Comparing Community Types with Normalized Mutual Information}

Because OSLOM detects \emph{coverings} rather than \emph{partitions} of users, standard cluster comparison methods like variation of information~\cite{meilua2003comparing} are not appropriate. Instead, we use a generalization of variation of information measure first introduced in~\cite{lancichinetti2009detecting}, the normalized information. \textbf{TK: Etc. Put more here about NMI, what it does, and how to interpret it.}

The normalized mutual information between the various community-types are shown in Figure~\ref{Fig-compare_coverings}. We see that similarity between the coverings is dictated by the generic covering type, with distinct community structure between the structural, activity-based, and interaction-based communities. Interestingly, the communities resulting from the different weightings are all more similar to each other than to the structural communities from the unweighted network.Also note that the communities based on the hashtag similarities are different from both the activity-based and mention-retweet-based communities.

\begin{figure}[h!]
  \centering
\includegraphics[width=0.50\textwidth]{figures/nmi_singletons.pdf}
\caption{The normalized mutual information between the communities using the different weightings. Weighting 0 corresponds to the structural (binary weighting) network, weightings 1 through 6 correspond to the weighting using the transfer entropy with lag 1 through 6, weighting 7 corresponds to the hashtag similarity, and weightings 8, 9, and 10 correspond to the mention, retweet, and mention-retweet weightings. Values of normalized mutual information close to 1 indicate similarity in the community structure, while values close to 0 indicate dissimilarity. The normalized mutual information is computed with the singletons removed.}
\label{Fig-compare_coverings}
\end{figure}

\subsection{Comparing Edges Across Different Community-Types}

We have defined communities using four different criteria: structure, activity, interaction, and content. For a fixed community type, edges for a particular community may be partitioned into three sets: those from a node in the community to another node in the community (internal-to-internal), those from a node in the community to a node outside of the community (internal-to-external), and those from a node outside the community to a node inside the community (external-to-internal). See Figure~\ref{Fig-edge_types} for a schematic of this edge partitioning. For a meaningful community, we expect the distribution of weights within the community (internal-to-internal weights) to be different from the distribution of weights without the community (internal-to-external and external-to-internal).

\begin{figure}[h!]
  \centering
\includegraphics[width=0.50\textwidth]{figures/edge-types.eps}
\caption{An example of the edges considered in determining the edge weight distribution for a given community (the focal community is in yellow). We focus on the internal-to-internal (red), internal-to-external (green), and external-to-internal (blue) edges. For a given focal community, all other edges (grey) are not considered.}
\label{Fig-edge_types}
\end{figure}

For example, Figure~\ref{Fig-edge_types} shows the distributions of hashtag-based weights for the largest community defined by the mention-retweet network. We see that the distribution of internal-to-internal hashtag weights has a longer tail than either the external-to-internal or internal-to-external hashtag weights, with edges within the community having higher weights than edges crossing the boundary of the community. Thus, while the community was defined in terms of interactions, we still see a shift in the distribution of \emph{topic-similarity}.

\begin{figure}[h!]
  \centering
\includegraphics[width=0.50\textwidth]{Figures/comm0_labels-mention-retweet_weights-hashtag-ecdf.pdf}
\caption{The proportion of edges with a weight at least as large as the weight on the horizontal axis, across the types of edges described in Figure~\ref{Fig-edge_types}. The community is defined by user interactions, and the edge weights are determined by topics.}
\label{Fig-dist_across_community}
\end{figure}

This shift in the distribution was typical of many of the community type / weight pairings. To explore this further, for each of the top 100 largest communities defined by a particular community type (structure-based, behavior-based, content-based, or activity-based), we computed the median internal-to-internal, external-to-internal, and internal-to-external weight for that community for each weight type. These values summarize the typical value of each of these types of edges. We can then compute the ratio of the median internal-to-internal weight and the median external/internal-to-internal/external weight, which captures the proportional shift in the distribution. Finally, we compute the median of this proportional shift across all 100 of the largest communities. These values are summarized in Table~\ref{Table-medians}.

% \begin{table}
% 	\caption{The median value for the ratio of the median external/internal-to-internal/external weight to median internal-to-internal weight for the different community/weight pairings.}
% 	\centering
% 	\begin{tabular}{c | c | c  c  c  c}
% 		& & \multicolumn{4}{ c }{Community} \\ \hline
% 		\multirow{4}{*}{Weight} & & S & TE & MR & HT \\ \hline
% 		& TE & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$\\
% 		& MR & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$\\
% 		& HT & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$ & $0.0 / 0.0$
% 	\end{tabular}
% \end{table}

\begin{table}
	\label{Table-medians}
	\caption{The median value across the 100 largest communities for the ratio of the median internal-to-internal weight to median external/internal-to-internal/external weight for the different community/weight pairings. \\ * For mention-retweet weights, weight zero edges were excluded from the computation of the median.}
	\centering
	\begin{tabular}{c | c | c  c  c}
		& & \multicolumn{3}{ c }{Weight} \\ \hline
		\multirow{4}{*}{Community} & & TE & MR & HT \\ \hline
		& S & $0.96/0.94$& $1.7/2.1$*& $9.0/8.0$\\
		& TE & $1.0/0.96$& $1.5/2.4$*& $24/17$\\
		& MR & $0.83/0.86$& $3.2/4.4$& $10/8.5$\\
		& HT & $0.9/0.89$& $2.4/2.6$*& $28/26$
	\end{tabular}
\end{table}

We see that overall, the