### 2014-01-14 paper/paper.tex

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{amssymb, amsmath, amsthm}
\usepackage{graphicx}
\usepackage{multirow}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Input Your Paper Title Here)
/Author (John Doe, Jane Doe)
/Keywords (Input your paper keywords in this optional area)}

%%%%%%%%%%
% Section Numbers
% Uncomment if you want to use section numbers % and change the 0 to a 1 or 2
\setcounter{secnumdepth}{0}

%%%%%%%%%%
% Body of Paper Begins
\begin{document}

% Title, Author, and Address Information
\title{
%Followers are not enough: Beyond structure in the detection of communities \\
Followers are Not Enough: Beyond Structural Communities in Online Social Networks \\
%Followers are not enough: Activity, Interaction, and Topic-based Communities in Online Social Networks \\
%Followers are not enough: Beyond structural communities in online social networks
% \\
 %Followers are not enough: exploring communities beyond structure
 }
%   \author{David Darmon  \\
%  University of Maryland\\
% Dept. of Mathematics \\
% ddarmon@math.umd.edu \\
%  \And Elisa Omodei \\
%  LaTTiCe (CNRS, ENS, Paris 3)\\
% ISC-PIF\\
% elisa.omodei@ens.fr \\
%  \And 
%  Joshua Garland\\
%  University of Colorado at Boulder\\
% Dept. of Computer Science \\
% joshua.garland@colorado.edu}

\maketitle

\begin{abstract}
\begin{quote}

[[Joshua]]

[[David]]

[[Elisa]]

\end{quote}
\end{abstract}

\input{parts/intro.tex}
\input{parts/community.tex}
\input{parts/methods.tex}
\input{parts/results.tex}
\input{parts/discussion.tex}
\input{parts/conclusions.tex}

\bibliography{references}
\bibliographystyle{aaai}

\end{document}